Overview and Summary of the First Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE)

### 2014-02-12 WSSSPE1_report/WSSSPE1_summary_report.tex
Mainly Section 3: Keynote and also other sections: fixed typos, added references.

The first keynote~\cite{WSSSPE1-keynote1} was delivered by Philip
E. Bourne of University of California, San Diego.  Bourne is a basic
biomedical scientist but has formed four software companies. He also
helped found PLOS and PDB~\cite{pdb-web}.  He is working on automating
three-dimensional visualizations of cell contents and molecular
structures, a problem that has not been solved and when done, would
serve as a key function of software in biomedical sciences.

Bourne's keynote presentation was entitled ``A Recipe for Sustainable
Software.''  He emphasized that sustainability for software ``does not just
mean more money from Government.''  He noted that in places, development in
science has improved thanks to open source and Github, but for the most part
remains arcane. He argued that there is much we can learn from the App Store
model. He mentioned BioJava~\cite{biojava-web} and
OpenScienceDataCloud~\cite{osdc-web} as distribution channels.  On a related
note, Bourne observed a common evolutionary pathway for computational biology
projects, from data archive to analytics platform to educational use, and
suggested that use of scientific software for outreach might be the final step.

@@

He also had problems with selling research software: the technology transfer
office in his university wanted huge IP reachthrough, whereby they would get a
share of profits from drugs developed by pharma companies who use the software.
He was aware this was unrealistic but the technology transfer office insisted
for a while. He wants to push a one-click approach to purchasing
university-written software.

@@

On the idea of the \textbf{digital enterprise}, where IT underpins the whole of
organizational activities, he contended that universities are way behind the
curve on this. In particular, he highlighted the separation of research,
teaching and administration into silos without a common IT framework as a
blocker to many useful organizational innovations: ``University 2.0 is yet to
happen.'' He spoke of a circumstance where someone had used an algorithm
developed for computational biology in marketing.

@@

Smith's talk started with an example from data reduction in Astronomy, where he
needed to remove interfering effects from the device. He built a ``bad
pixel mask'', and realized that while it was persistent, there was no
way/practice of sharing this with the data. Consequently many researchers
repeated the same calculations. He estimated that 13 person-years were
wasted in this redundant calculation.

@@

In the opening keynote, Phil Bourne suggested that perhaps sustainability can be defined
as the effort that happens to make the essential things continue. This leads to
having to decide what it is that we want to sustain, if what we want to sustain is
valuable, and finally, who would care if it went away, and how does one measure
how much they care.

@@

\begin{itemize}
    \item github
    \item mvapich
    \item Cactus
    \item visTrails
    \item MINRES/QLP
    \item SimGrid
    \item Swift
    \item Galaxy
    \item FastMath and package managers
    \item R/qtl
    \item OWL/RDF
    \item AMR/multiphysics
    \item DUNE
    \item PAAS
    \item Gateways
    \item Globus-Genomics
\end{itemize}

@@

--Science Oriented
      Software developed to solve domain science problems eg. MINRES/QLP

>>>>

The first keynote~\cite{WSSSPE1-keynote1} was delivered by Philip
E. Bourne of University of California, San Diego.  Bourne is a basic
biomedical scientist but has formed four software companies. He also
helped found PLOS~\cite{plos-web} and PDB~\cite{pdb-web}.  He is working on automating
three-dimensional visualizations of cell contents and molecular
structures, a problem that has not been solved and when done, would
serve as a key function of software in biomedical sciences.

Bourne's keynote presentation was entitled ``A Recipe for Sustainable
Software.''  He emphasized that sustainability for software ``does not just
mean more money from Government.''  He noted that in places, development in
science has improved thanks to open source and GitHub~\cite{github-web}, but for the most part
remains arcane. He argued that there is much we can learn from the App Store
model. He mentioned BioJava~\cite{biojava-web} and
Open Science Data Cloud~\cite{osdc-web} as distribution channels.  On a related
note, Bourne observed a common evolutionary pathway for computational biology
projects, from data archive to analytics platform to educational use, and
suggested that use of scientific software for outreach might be the final step.

@@

He also had problems with selling research software: the technology transfer
office in his university wanted huge IP reach through, whereby they would get a
share of profits from drugs developed by pharmaceutical companies who use the software.
He was aware this was unrealistic but the technology transfer office insisted
for a while. He wants to push a one-click approach to purchasing
university-written software.

@@

On the idea of the \emph{digital enterprise}, where IT underpins the whole of
organizational activities, he contended that universities are way behind the
curve on this. In particular, he highlighted the separation of research,
teaching and administration into silos without a common IT framework as a
blocker to many useful organizational innovations: ``University 2.0 is yet to
happen.'' He spoke of a circumstance where someone had used an algorithm
developed for computational biology in marketing.

@@

Smith's talk started with an example from data reduction in Astronomy, where he
needed to remove interfering effects from the device. He built a ``bad
pixel mask,'' and realized that while it was persistent, there was no
way/practice of sharing this with the data. Consequently many researchers
repeated the same calculations. He estimated that 13 person-years were
wasted in this redundant calculation.

@@

In the opening keynote, Phil Bourne suggested that perhaps sustainability can be defined
as the effort that happens to make the essential things continue. This leads to
having to decide what it is that we want to sustain, whether what we want to sustain is
valuable, and finally, who would care if it went away, and how does one measure
how much they care.

@@

\begin{itemize}
    \item github
    \item mvapich
    \item Cactus
    \item visTrails
    \item MINRES-QLP
    \item SimGrid
    \item Swift
    \item Galaxy
    \item FastMath and package managers
    \item R/qtl
    \item OWL/RDF
    \item AMR/multiphysics
    \item DUNE
    \item PAAS
    \item Gateways
    \item Globus-Genomics
\end{itemize}

@@

--Science Oriented
      Software developed to solve domain science problems eg. MINRES-QLP
